Subset coercions in \Coq

### 2006-09-14 papers/article/main.tex
CPM fixes

\documentclass{llncs}
\usepackage{abbrevs}
\usepackage{array}
\usepackage{alltt}
\usepackage{color}
\def\text{\mbox}
%\pagestyle{plain}
%\usepackage{amsmath}
%\usepackage{amsthm}

@@

\newenvironment{code}{\begin{alltt}}{\end{alltt}}
% Caml2Tex

\definecolor{KWLanguage_color}{rgb}{0.6,0,0}
\definecolor{KWStructure_color}{rgb}{0.8,0.8,0}
\definecolor{KWModule_color}{rgb}{0.7,0.5,0}
\definecolor{KWSubprog_color}{rgb}{0.6,0,0.6}
\definecolor{KWException_color}{rgb}{0.8,0,0}
\definecolor{Constructor_color}{rgb}{0,0.5,0}
\definecolor{Comment_color}{rgb}{0.6,0.6,0.6}
\definecolor{String_color}{rgb}{0.6,0.5,0.5}
\definecolor{Varfunction_color}{rgb}{0.2,0.2,0.8}
\definecolor{Constant_color}{rgb}{0.6,0.5,0.5}
\definecolor{Variable_color}{rgb}{0.2,0.2,0.8}

\newcommand\KWLanguage[1]{\rmfamily{\color{KWLanguage_color}{#1}}}
\newcommand\KWStructure[1]{\sffamily{\color{KWStructure_color}{#1}}}
\newcommand\KWModule[1]{\rmfamily{\color{KWModule_color}{#1}}}
\newcommand\KWSubprog[1]{\rmfamily{\color{KWSubprog_color}{#1}}}
\newcommand\KWException[1]{\rmfamily{\color{KWException_color}{#1}}}
\newcommand\Constructor[1]{\textbf{\color{Constructor_color}{#1}}}
\newcommand\Comment[1]{\rmfamily{\color{Comment_color}{#1}}}
\newcommand\String[1]{\rmfamily{\color{String_color}{#1}}}
\newcommand\Varfunction[1]{\sffamily{\color{Varfunction_color}{#1}}}
\newcommand\AVariable[1]{\textit{\color{Variable_color}{#1}}}
\newcommand\Constant[1]{\texttt{\color{Constant_color}{#1}}}
\newcommand\pow{$^{\wedge}$}
\newcommand\bs{$\backslash$}

@@

\institute{LRI, Paris Sud XI University\\\email{sozeau@lri.fr}}
\title{Subset coercions in \Coq \\ \small{Draft - \today}}
\date{\today}

@@

\section{Implementation}
\label{sec:Program}
This mechanism has been implemented in \Coq and tested on simple
examples. We actually generate coercions simultaneously with typing 
in the implementation, which is sound thanks to the previous proofs. Our
typing is a clone of \Coq's original typing algorithm, so we benefit
from all the features of \Coq including implicit variables, notations and the existing
coercion system. The type-checker also transforms pattern-matching so
that an equality between the matched term and the pattern is present in
each branch typing context and it allows the user to put explicit holes in the term, for example:
\begin{code}
  Program Definition hd ( l : list A | l <> [] ) : A :=
    match l with
    | hd :: tl -> hd
    | [] -> _
    end.
\end{code}
Here the second branch can't be taken because we would have $l = []$ and
$l \neq []$, but we may not have any placeholder element of type
$A$. There is also support for strucural and well-founded \verb|Fixpoint| definitions.

@@

\subsection{Related work}
Other efforts to build certified programs using type theory include
Ca\-yenne \cite{Augustsson99} and Epigram
\cite{DBLP:journals/jfp/McBrideM04}. Cayenne offers dependent
types and general recursion at the expense of a non-terminating type
checker; it can be thought of as a testbed for developing languages with
dependent types but is not so much aimed at building certified programs.
Epigram ought to become a complete programming language with dependent
types at its heart. Instead of using a phase distinction to separate
coding and proving, Epigram is based on an interactive, 2-dimensional editing process
where code and proofs are written incrementally to get a complete program.
Using type annotations and with the help of the editor for structuring, 
one is able to write programs with precise specifications. However it
is not obvious if it could be made usable by programmers because of the
paradigm change and scaling issues. DML \cite{Xi98:PhD} is more akin to
comparison with our solution. It provides a way to use a restricted set
of dependent types in \ML programs, carefully chosen so as to keep
type-checking decidable with the help of an automatic prover. Our method
subsumes this one as it is perfectly possible to integrate automatic
proof tools with \Coq to discharge the generated obligations.

>>>>

\documentclass{llncs}
\usepackage{abbrevs}
\usepackage{array}
\usepackage{alltt}
\usepackage{color}
\def\text{\mbox}
\pagestyle{plain}
%\usepackage{amsmath}
%\usepackage{amsthm}

@@

\newenvironment{code}{\begin{alltt}}{\end{alltt}}

@@

\institute{LRI, Paris Sud XI University\\\email{sozeau@lri.fr}}
\title{Subset coercions in \Coq} % \\ \small{Draft - \today}}
\date{\today}

@@

\section{Implementation}
\label{sec:Program}
This mechanism has been implemented in \Coq and tested on simple
examples. We actually generate coercions simultaneously with typing 
in the implementation, which is sound thanks to the previous proofs. Our
typing is a clone of \Coq's original typing algorithm, so we benefit
from all the features of \Coq including implicit variables, notations and the existing
coercion system. The prototype also contains support for strucural and 
well-founded \verb|Fixpoint| definitions.
The type-checker transforms pattern-matching so
that an equality between the matched term and the pattern is present in
the typing context of each branch and it allows the user to put explicit 
holes in the term, for example:
\begin{code}
  Program Definition hd ( l : list A | l <> [] ) : A :=
    match l with
    | hd :: tl -> hd
    | [] -> _
    end.
\end{code}
Here the second branch can't be taken because we would have $l = []$ and
$l \neq []$, but we may not have any placeholder element of type
$A$.

It must be stressed that we do not lose any information when
generating obligations as we have the whole term context at hand,
avoiding the complicated proof obligation generation of \PVS.
Here we would have $l : \mysubset{l}{\texttt{list}~A}{l \neq []}, Heql : \eltpit~l =
[]$. We have tactics that clean the goal and context (deconstructing
subsets and simplifying to present $l : \texttt{list}~A, H : l \neq [], Heql : l = []$
to the user) but it does not match the usability of \PVS yet.

@@

\subsection{Related work}
Other efforts to build certified programs using type theory include
Ca\-yenne \cite{Augustsson99} and Epigram
\cite{DBLP:journals/jfp/McBrideM04}. Cayenne offers dependent
types and general recursion at the expense of a non-terminating type
checker; it can be thought of as a testbed for developing languages with
dependent types but is not so much aimed at building certified programs.
Epigram ought to become a complete programming language with dependent
types at its heart. Instead of using a phase distinction to separate
coding and proving, Epigram is based on an interactive, 2-dimensional editing process
where code and proofs are written incrementally to get a complete program.
Using type annotations and with the help of the editor for structuring, 
one is able to write programs with precis e specifications. However it
is not obvious if it could be made usable by programmers because of the
paradigm change and scaling issues. DML \cite{Xi98:PhD} is more akin to
comparison with our solution. It provides a way to use a restricted set
of dependent types in \ML programs, carefully chosen so as to keep
type-checking decidable with the help of an automatic prover. Our method
subsumes this one as it is perfectly possible to integrate automatic
proof tools with \Coq to discharge the generated obligations.
